\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsthm}

\usepackage[T1]{fontenc}
\linespread{1.05}         % Palatino needs more leading (space between lines)
\usepackage[hmargin=1.0in]{geometry}

\bibliographystyle{apalike}

\begin{document}

\title{Order conditions for multistep Runge-Kutta methods}
\author{David I. Ketcheson} 
\maketitle

\input{macros}

\section{Review of Order Conditions for Runge--Kutta Methods}
  \input{rkoc}

\section{Enumeration of conditions}
  \input{rkocenum}

\section{Multistep Runge--Kutta Methods of Type I}
  \input{mrkoc}

\section{Spijker forms}
The Spijker form for the Type I methods of the last section is
\begin{align}
\bx & = \left[u^{n-k+1},u^{n-k+2},\dots,u^{n-1},u^n\right] \\
\by & = \left[u^{n-k+1},u^{n-k+2},\dots,u^{n-1},u^n,y_1,\dots,y_s,u^{n+1}\right] \\
\mS & = \begin{pmatrix}\mI\\\mD\\\btheta\transpose\end{pmatrix} \ \ \ \
\mT   = \begin{pmatrix}\mzero & \mzero & 0 \\ \mzero & \mA & \mzero \\ \mzero & \bb\transpose & 0 \end{pmatrix}.
\end{align}

We are also interested in Type II methods, which take the form
\begin{subequations} \label{eq:mrk-typeII}
\begin{align}
y_i^n & = \sum_{l=1}^{k} d_{il} u^{n-k+l} + \Dt\sum_{l=1}^{k-1} \hat{a}_{il} F(u^{n-k+l}) + \Dt\sum_{j=1}^s a_{ij} F(y_j^n) \\
u^{n+1} & = \sum_{l=1}^{k} \theta_l u^{n-k+l} + \Dt\sum_{l=1}^{k-1} \hat{b}_{l} F(u^{n-k+l}) + \Dt\sum_{j=1}^s b_j F(y_j^n),
\end{align}
\end{subequations}
where $y_1^n=u^n$.  These admit the Spijker form
\begin{align}
\bx & = \left[u^{n-k+1},u^{n-k+2},\dots,u^{n-1},u^n\right] \\
\by & = \left[u^{n-k+1},u^{n-k+2},\dots,u^{n-1},u^n,y_1=u^{n},y_2,\dots,y_s,u^{n+1}\right] \\
\mS & = \begin{pmatrix}\mI \ \mzero \\ \mD \\\btheta\transpose \end{pmatrix} \ \ \ \
\mT   = \begin{pmatrix}\mzero & \mzero & 0 \\ \mAh & \mA & \mzero \\ \bbh\transpose & \bb\transpose & 0 \end{pmatrix}.
\end{align}
Here the first row of $\mD$ is $(0,0,\dots,0,1)$ and the first row of
$\mA,\mAh$ is identically zero.
The order conditions presented above can also be applied to these Type II methods
by writing them as Type I methods (simply introduce additional stages that are
equal to the old steps).  {\bf David can write this last bit up if necessary.}

\section{Methods for linear problems}
In this section we consider only linear problems of the form $$u_t = L u,$$
so that the order conditions simplify considerably.  This case is of 
interest for several reasons.  First, linear problems are of interest in many 
cases (cite Bernardo Cockburn), and it is possible to include a time-dependent 
forcing term or boundary condition while maintaining the order of the method. 
Second, up to second order  the order conditions for  linear and nonlinear
problems are the same.  Thus, this approach will provide results for second 
order multi-step multi-stage methods, which gives us a picture of the size 
of the time-step for higher order nonlinear methods.  Finally, the time-step 
restriction for the linear high order methods serves as an upper bound for 
that of nonlinear high order methods. 

\subsection{Tall Tree Order conditions for Multistep Runge--Kutta methods}
For $k$-step Runge-Kutta methods of Type II,
applying the method to the linear scalar homogeneous test equation, we
find that it reduces to the iteration
\be
u^{n+1} = \left( \btheta\transpose + z \hat{\vb}\transpose + z \vb\transpose ( I-z\mA)^{-1} (\mD+z\hat{\mA}) \right) \vu
\ee
where
\be
\vu = \left(u^{n-k+1},u^{n-k+2},\dots,u^{n}\right)^T
\ee
and $\mAh,\vbh\transpose$ each have $k$ columns but the last column is zero.
Using the Taylor expansions 
\begin{subequations}
\begin{align}
\left( \mI - z\mA \right)^{-1}  & = \sum_{j=0}^\infty z^j \mA^j \\
u^{n-\vl} = e^{-\vl\Dt}u^n  & = u^n \sum_{j=0}^\infty \frac{1}{j!} z^j (-l)^j
\end{align}
\end{subequations}
we obtain the tall tree order conditions for linear problems.
These are given by equating coefficients in
\begin{align}
e^z + \Oop(z^{p+1}) & = 
    \left[ \btheta\transpose + z \bhT + z \bT
    \left(\sum_{j=0}^\infty z^j \mA^j \right) (z + \mD \hat{\mA})\right]
    \left(\sum_{j=0}^\infty \frac{1}{j!}z^j (-\vl)^j\right) \nonumber \\
 & = \sum_{j=0}^\infty z^j {\huge(} \frac{(-1)^j}{j!} \btheta\transpose \vl^j
    + \frac{(-1)^{j-1}}{(j-1)!}\bhT \vl^{j-1} \\
    & + \bT \sum_{i=0}^{j-1}\frac{(-1)^i}{i!}\mA^{j-i-1}\mD\vl^i
    + \frac{(-1)^{i-1}}{(i-1)!}\mA^{j-i}\hat{\mA}\vl^{i-1} \nonumber
    {\huge)}
\end{align}
where $\vl=(k-1,k-2,\dots,0)\transpose$.
{\em need to add something about terms with negative exponents
being understood to be zero here}

The resulting conditions are (given our assumption that $\mD\ve=\ve$ and 
$\btheta\transpose\ve=1$):
\begin{subequations}
\begin{align}
(\bT + \bhT)\ve - \btheta\transpose\vl & = 1 \\
\bT\vc - \bhT\vl
    + \frac{1}{2}\btheta\transpose\vl^2 & = \frac{1}{2} \\
\bT(\mA\vc + \frac{1}{2}\mD\vl^2 - \mAh\vl) +\frac{1}{2}\bhT\vl^2 
    - \frac{1}{6}\btheta\transpose\vl^3 & = \frac{1}{6}
\end{align}
\end{subequations}
where
$$ \vc = (\mA+\hat{\mA})\ve - \mD\vl.$$


For $k$-step Runge-Kutta methods of Type I,
the analysis is the same as that above, but with
$\hat{\vb}=\mAh=0$.
The first five order conditions in this case reduce to
\begin{subequations}
\label{2ndOCs}
\begin{align}
\vb^T \ve - \btheta\transpose \vl &= 1, \\
\vb^T \vc + \frac{1}{2} \btheta\transpose \vl^2 & = \frac{1}{2}, \\
\vb\transpose \left(\mA\vc + \frac{1}{2}\mD\vl^2\right) - \frac{1}{6} \btheta\transpose\vl^3 & = \frac{1}{6} \\
\vb\transpose \left(\mA^2\vc + \frac{1}{2}\mA\mD\vl^2 - \frac{1}{6}\mD\vl^3\right) + \frac{1}{24} \btheta\transpose\vl^4 & = \frac{1}{24} \\
\vb\transpose \left(\mA^3\vc + \frac{1}{2}\mA^2\mD\vl^2 - \frac{1}{6}\mA\mD\vl^3 + \frac{1}{24}\mD\vl^4\right) - \frac{1}{120} \btheta\transpose\vl^5 & = \frac{1}{120}
\end{align}
\end{subequations}
where $\vc=\mA\ve-\mD\vl$.


\subsection{Optimal SSP Multistep Runge--Kutta methods for linear problems}

{\bf These are David's old results from 2007-2008.}
SSP coefficients for optimal explicit methods of Type II are shown in Tables
\ref{tbl-2order}-\ref{tbl-3order}.
We have solved numerically the optimization problem of maximizing $r$ subject to
(\ref{2ndOCs}) and (\ref{spijkermon}) for various values of $s,k$ with $p=2$
(i.e., implicit methods of Type I).
It appears (after some \verb fmincon  searches) that the optimal 2nd order 
SSP implicit Runge-Kutta methods given in \cite{ferracina2008,ketcheson2008b} 
are actually optimal over all second order $k$-step, 2-stage methods.

\begin{table}
\center
\begin{tabular}{l|cccc} \hline
s \ k & 2     & 3     & 4     & 5 \\ \hline
2 & 0.707 & 0.809 & 0.860 &  \\
3 & 0.817 & 0.879 & 0.911* & \\
4 & 0.866 & 0.911 & 0.934 & \\
5 & 0.894 & 0.930 &  & \\  \hline
\end{tabular}
\caption{Effective SSP coefficients $\ceff$ of optimal
  explicit 2nd order $k$-step Runge-Kutta methods of type (explicit with
  a zero row) (for both linear and nonlinear problems).}
\label{tbl-2order}
\end{table}


\begin{table}
\center
\begin{tabular}{l|cccc} \hline
s \ k & 2     & 3     & 4     & 5 \\ \hline
2 & 0.366 & 0.556 & 0.622 & 0.622 \\
3 & 0.550 & 0.667*& 0.667 & \\
4 & 0.627 &       &       & \\
5 & 0.677 & 0.679 & 0.679 & \\ 
6 & 0.705 & 0.705 & 0.705      & \\ \hline
\end{tabular}
\caption{Effective SSP coefficients $\ceff$ of optimal
  explicit 3rd order $k$-step Runge-Kutta methods of type (explicit with
  a zero row) for linear problems.}
\label{tbl-3order}
\end{table}



\bibliography{rk_order_conditions}

\end{document}
