\documentclass[12pt]{article}

\usepackage[T1]{fontenc}
\linespread{1.05}         % Palatino needs more leading (space between lines)
\usepackage[hmargin=1.0in]{geometry}

\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsthm}

\bibliographystyle{apalike}

\begin{document}

\title{Order conditions for Runge-Kutta methods}
\author{David I. Ketcheson} 
\maketitle

\input{macros}

A Runge--Kutta method 
$$
u^{n+1} = u^n + \Delta t \sum_j b_j f(y_j, t^n + c_j \Dt)
$$
is similar to a quadrature rule with weights $b_j$ and
abscissas $c_j$.  But the Runge--Kutta method is more complex, since 
intermediate approximations must be performed to obtain the stage values
$y_j$, meaning that the values $f(y_j)$ appearing in the quadrature formula
are not exact derivatives of the
solution.  The order conditions for Runge--Kutta methods are precisely 
the necessary conditions to ensure that the resulting errors in the final
approximation vanish up to a given order in $\Delta t$.

\section{Derivation of the order conditions\label{sec:rkoc}}
  \input{rkoc}

\section{Enumeration of conditions}
  \input{rkocenum}

\section{Order conditions for dense output}
Let us consider now continuous Runge--Kutta methods, which are equipped
with an additional set of weights $b_j(\theta)$ such that
\begin{align} \label{dense-output}
    u^{n+\theta} & = u^n + \Dt \sum_{j=1}^s b_j(\theta) f(y_j) \approx u(t_n + \theta \Dt).
\end{align}
Here we are generally interested in values $0\le \theta \le 1$, i.e.\ the
$\theta$-dependent weights are used to interpolate and provide output
at an arbitrary time within a given step.  Formula \eqref{dense-output}
is known as a {\em dense output} formula.  It avoids the need to ensure
that the integration hits desired output times exactly.  Most importantly,
it ensures that step sizes are determined only by accuracy and stability
considerations, even if very frequent output is required.

How accurate should the approximation \eqref{dense-output} be?
At first glance, it seems like the dense output formula ought
to have the same accuracy as that given by the regular weights (i.e., $p$).
However, since the dense output will not be used to propagate the solution,
and hence contributes only to local (not global) error, it is sufficient
to require order $p-1$.

What are the order conditions on the weights $b_j(\theta)$?  Here again
we will follow Albrecht's approach.  The standard references use Butcher's
approach and find that all of the order conditions become $\theta$-dependent.
But with Albrecht's approach we will see that only the bushy-tree conditions
depend on $\theta$.

To derive the order conditions, we follow the analysis of Section~\ref{sec:rkoc},
but in place of $u^{n+1}$ we consider the approximation $u^{n+\theta}$ above.
We find that 
\begin{align*}
\tu(t_{n+\theta}) & = \tu^{n} + \Dt \bb(\theta)\transpose\btf^n + \Dt \lte^n(\theta)
\end{align*}
where
\begin{align*}
\lte^n(\theta) & = \sum_{k=1}^\infty \lte_k \Dt^{k-1} \tu^{(k)}(t_{n})
\end{align*}
where
\begin{align*}
\lte_k & = \frac{\theta^k}{k!} - \frac{1}{(k-1)!} \bb\transpose \bc^{k-1}
\end{align*}
Notice that everything but the $\lte_k$ is unchanged from our earlier
analysis.  Thus the order conditions for order $p$ are
\begin{align*}    
\bb(\theta)\transpose \bc^{k-1} & = \frac{\theta^k}{k}  &  \mbox{for }  & 0\le k\le p \\
\bb(\theta)\transpose\rhsserr^n_k & = 0 &  \mbox{for } & 0\le k\le p-1.
\end{align*}
As predicted, only the bushy-tree conditions are modified.
Thus the vector $\bb(\theta)$ must lie, for every $\theta$, in the same
subspace determined by orthogonality to the vectors $\rhsserr_k$.


\bibliography{rk_order_conditions}

\end{document}
